# Our case
  
  Our study investigates the diachronic shift in the usage of the two Japanese quasi-synonyms 自殺 (jisatsu, “to commit suicide”) and 自害 (jigai, “to commit suicide, often in a ritual or formal context”).
  Representative examples
In Natsume Sōseki’s Meiji-period short story we find:

彼は病苦に耐えかねて自殺を遂げた。
(He could no longer bear the suffering of his illness and committed suicide.)

武士は名誉を守るために自害を決行した。
(The samurai carried out suicide in order to preserve his honor.)

In a Shōwa-era novella by Akutagawa Ryūnosuke:

その若者は深い絶望の中で自殺を試み、
(That young man attempted suicide in the depths of despair,)

戦場の勇士は誇り高く自害した。
(and the brave warrior on the battlefield proudly carried out suicide.)

自殺 (jisatsu)
自らの命を絶つこと。自分の意思で生命を終える行為。
(To take one’s own life; an act of ending one’s life by one’s own volition.)
転じて、自己否定的・絶望的行為全般を指す。
(By extension, refers to any self-negating or despair-driven act.)

自害 (jigai)
自分に害（自ら命を絶つ）を及ぼすこと。特に武士の切腹など儀礼的・形式的自死を含む。
(To inflict harm on oneself—i.e., to end one’s own life—especially including ritual or formal suicide such as a samurai’s seppuku.)
自分を犠牲にして目的を果たす行為や強い決意を示す文脈で用いられる。
(Used in contexts indicating an act of self-sacrifice to achieve an objective or to show strong resolve.)

Explanation of term choice
In Japanese, there is also 切腹 (seppuku)—the ritual disembowelment historically performed by samurai—as a counterpart to “suicide.” However, we chose 自害 (jigai) for our comparison with 自殺 (jisatsu) for several reasons:

Broader semantic scope

切腹 (seppuku) refers very specifically to the formal samurai practice of self-disembowelment, bound by historical and class-based ritual contexts.

自害 (jigai), by contrast, appears not only in descriptions of ritual suicide but also in a wider range of literary and historical contexts—covering any self-inflicted death carried out with a sense of solemnity or duty, beyond strictly samurai circles.

Diachronic prevalence

In pre-Meiji and Meiji-period texts, 自害 is the dominant literary term for formal or honor-driven self-harm, whereas 切腹 is reserved almost exclusively for narratives involving samurai.

From Taisho onward, 自害 continues to surface in modern fiction and non-fiction whenever an author wishes to evoke a formal or dignified self-sacrifice, while 切腹 remains tied to historical or military settings.

Quasi-synonymic contrast

Because our goal is to trace the gradual shift from any formalized self-harm (自害) to the more clinical or general term 自殺, 自害 provides the best “paired” counterpart: it shares the core sense of “self-caused death” yet occupies a distinct register and pragmatic function.

Thus, by comparing 自殺 and 自害, we capture the evolution from a broad literary usage of formalized self-harm to the modern, neutral designation of suicide—something that 切腹 alone could not illustrate.



# Our data

  We drew a balanced corpus of 500 texts from Aozora Bunko, stratified across the Meiji, Taisho, Showa, Heisei and Reiwa eras: 250 texts containing 自害 and 250 texts containing 自殺. All texts were preprocessed to remove Aozora editorial markers, ruby-notation explanations, and introductory metadata, then flattened into a single line to facilitate context extraction.
  (разметка - ворк ин прогресс, по задумке - ручная разметка + обучить на ручной разметке TF-IDF + LogisticRegression)
  
  
# Hypothesis
(изначальные гипотезы - диахронические. Можно ли их и оставить? Будет ли этого достаточно для исследования?)
**The null hypothesis:** The average publication year for contexts containing 自殺 (jisatsu) is equal to the average publication year for those containing 自害 (jigai).  
**The alternative hypothesis:** The average publication years for the two groups are different.

```{r include = FALSE}
library(tidyverse)
library(TTR)
library(tseries)
library(forecast)
```

(вторые гипотезы на случай, если этого окажется мало, а его будто бы кажется)
**The null hypothesis:** The proportion of ritual contexts is the same in the 自殺 group and the 自害 group.
**The alternative hypothesis:**The proportion of ritual contexts is higher for 自害 than for 自殺.
# Data

```{r}
csv_url <- 'https://raw.githubusercontent.com/nightmarewarrior/data_for_R/refs/heads/main/dataset_context_with_rule.csv'
df <- read_csv(csv_url)
df
```

Let's take a look at the diagrams which can show the tendencies of usage of these verbs to make some empirical conclusions. Our distributions are:

```{r}
summary_tbl <- df %>%
  count(era, keyword) %>%
  arrange(era, keyword)
print(summary_tbl)
```
```{r}
ggplot(summary_tbl, aes(x = era, y = n, fill = keyword)) +
  geom_col(position = "dodge") +
  labs(
    x    = "Эпоха",
    y    = "Число контекстов",
    fill = "Термин",
    title= "Распределение контекстов 自害 и 自殺 по эпохам"
  ) +
  theme_minimal()
```
(перекос ожидаем - датасет требует балансировки, поскольку большинство оцифрованной литературы - эпохи Мэйдзи, и к тому же они длиннее. Возможно стоит сделать минидатасет в рамках этой эпохи?)


```{r}
library(Hmisc)
summary_df <- df %>%
  group_by(keyword) %>%
  summarise(
    mean_year = mean(publication_year),
    ci_low    = t.test(publication_year)$conf.int[1],
    ci_high   = t.test(publication_year)$conf.int[2]
  )
ggplot(summary_df, aes(x = keyword, y = mean_year, fill = keyword)) +
  geom_col() +
  geom_errorbar(aes(ymin = ci_low, ymax = ci_high), width = 0.2) +
  theme_minimal()
```



```{r}
merged_distr_sep <- rbind(mignut_init_distr, podmignut_init_distr)

merged_distr_sep %>%
  ggplot(aes(Publ_year, dist_per))+
  geom_smooth()+
  facet_wrap(~category)
```

Taking first look on these diagrams we can observe that our hypothesis actually tends to be true. 

But these are just some emprircal conclusions. Next thing that we did is applying some statistical tests to our data


# Applying statistical methods


```{r}
t.test(publication_year ~ keyword,
       data = df,
       var.equal = TRUE)
```
The obtained p-value is enough to proove our hipothesis, but let's think about our data in another way. Now let's talk about our values from point of view of the time series.

# Time series



```{r}
library(dplyr)
yearly_counts <- df %>%
  mutate(year = as.integer(substr(pub_date, 1, 4))) %>%
  count(year, keyword) %>%
  pivot_wider(names_from = keyword, values_from = n, values_fill = 0)


```




```{r}
library(ggplot2)
ggplot(yearly_counts, aes(x = year)) +
  geom_line(aes(y = jisatsu, color = "jisatsu")) +
  geom_line(aes(y = jigai,  color = "jigai")) +
  labs(x="Year", y="Count of contexts", color="Term") +
  theme_minimal()

```



```{r}
library(changepoint)
cp_jis <- cpt.meanvar(yearly_counts$jisatsu, method="PELT")
cp_jig <- cpt.meanvar(yearly_counts$jigai,  method="PELT")
plot(cp_jis); plot(cp_jig)
```




# Results

As a result, we have received confirmation of our hypothesis, however, the database for an ideal statistical analysis is not enough, but for existing occurrences we have observed some significant correlation.
